---
title: "R Function Cheat Sheet"
author: "Shunkei Kakimoto"
output:
  html_document:
    number_sections: yes
    theme: flatly
    toc_float: yes
    toc: yes
    toc_depth: 3
geometry: margin=1in
---

```{r setup, include=FALSE}
library(knitr)
library(here)

# here::i_am("Dropbox/R-project/R-functions/GitControlled/vis/R-vis.rmd")

# opts_knit$set(root.dir = "")
# opts_knit$set(root.dir = here())

knitr::opts_chunk$set(
  echo = TRUE,
  eval = FALSE,
  cache = FALSE,
  comment = NA,
  message = FALSE,
  warning = FALSE,
  tidy = FALSE,
  cache.lazy = FALSE,
  #--- figure ---#
  dpi = 400,
  fig.width = 7.5,
  fig.height = 5,
  out.width = "750px",
  out.height = "500px"
)

# /*===== Basic Packages  =====*/
# /*---- Data Wrangling ----*/
library(data.table)
library(tidyverse)
library(DescTools)
library(maps)

# /*---- Visualization ----*/
library(RColorBrewer)
library(patchwork)
library(ggplot2)
library(ggthemes)
library(ggpubr)
library(viridis)
library(grid)
library(gridExtra)
library(GGally)

# /*---- Model Summary ----*/
library(stats)
library(modelsummary)
library(flextable)
library(officer)
library(officedown)
library(gt)

```

# Contents

## Topics
+ [**tmap package**](#tmap)
  * Read [this](https://bookdown.org/nicohahn/making_maps_with_r5/docs/tmap.html) 
  * Read [this](https://geocompr.robinlovelace.net/adv-map.html) 
  * Overlay raster layers and vector data
+ [**Descriptive Statistical: ggpairs()**](#CorrelationMap)
+ [**theme() and guide() of ggplot2**](#theme)
  * position of legend 
  * change size of points in legend
+ [**Change ggplot facet lavels**](#GgplotFacetLavels)
+ [**Modify the legend**](#ChangeLegendName)
+ [**Plot Raster with facet**](#RasterFascet)
+ [**Color command**](#color)
+ [**ggplot with year**](#year)

## Useful Functions

+ [**coord_equal()**](#coord_equal)

+ [**scale_fill_viridis_c()**](#scale_fill_viridis_c)
  * make the figure color differentiate more
  * you can also change the legend name with scale_fill_viridis_c()

+ [**theme_void()**](#theme_void)

+ [**theme_bw()**](#theme_bw)

+ [**lab()**](#lab)
  * modify the legend 

+ [**scale_fill_distiller(pattern = “Green”, direction =1)**](#scale_fill_distiller)
  * change how the appearance of color bar
  
+ [**facet_grid(type~variable,scale='free’)**](#AdjustScaleAxis)
  * adjust the scale in the axis of the figures

+ [**theme(legend. title = element_blank())**](#RemoveLegendTitle)
  * remove legend title

+ [**theme(legend.position = “none”)**](#HideLegend)
  * hide the entire legend to crerate ggplot with no legend

+ [**theme(panel.grid = element_line(color='transparent'))**](#Transparent)
  * change the color of grid lines

---


# Topics

## tmap {#tmap}

+ `tm_shape( ) `

```{r}
left_join(field,data[,.(unique_cell_id,opt_N)],by='unique_cell_id') %>%
    select(opt_N) %>%
    tm_shape() +
      tm_fill(col = "opt_N")
```

<br>

```{r}
tm_shape(bavaria) +
  tm_fill()

tm_shape(bavaria) +
  tm_borders()

tm_shape(bavaria) +
  tm_polygons()
```

Color differentiated point data \
+ Use `tm_symbols()` or `tm_dots`

```{r}
tm_shape(hp_bd_buffer_100km)+
  tm_fill(alpha=0.6) +
tm_shape(hp_bd) +
  tm_fill(col="lightblue") +
tm_shape(unique_tg_corelogic_sf) +
  tm_dots(col = "Ogallala", size = 0.001)


tm_shape(hp_bd_buffer_100km)+
  tm_borders(col = "black", lwd = 1.5) +
tm_shape(hp_bd) +
  tm_borders(col="blue", lwd = 1.5) +
tm_shape(unique_tg_corelogic_sf) +
  tm_symbols(col = "Ogallala", shape = 19, size = 0.001)
```

+ tamp with leaflet

```{r}
tmap_leaflet(
  tm_shape(ne_tract) + 
    tm_polygons(col="pop_density")
    # tm_fill(col="totalpop")
)
```


```{r}
tm_shape(Europe) +
  tm_polygons(col = "outcome",
              style ="cont", palette = "-RdYlGn")

tm_shape(Europe) +
  tm_polygons(col = "outcome",
              style ="cont", palette = "RdYlGn")
```


```{r}
ne_pop %>%
  filter(County == "Douglas County") %>%
  tm_shape(.) + 
    tm_polygons(col="pop_den_sqmiles", style="order", lwd=0.3) +
    tm_layout(legend.outside = TRUE) 
```

```{r}
# === title and so on === #
tm_shape(mn_adm2)+
  tm_polygons(
    col = "Total Population, 2021",
    title = "Population",
    n = 8,
    style = "jenks"
  )+
  tm_layout(
      main.title = "2021 County−Level Population Estimates",
    legend.outside = TRUE
    )
```

### facets

```{r}
tm_shape(hp_counties)+
  tm_borders()+
tm_shape(hp_states)+
  tm_borders(col="blue")+
tm_shape(filter(nonag_cdl_land_sf, Crop %in% target_lu)) +
  tm_symbols(col="green", shape = 19, size=0.1)+
  tm_facets(by=c("Crop"), free.scales=FALSE)
```


### Overlay raster layers and vector data

```{r}
tm_shape(masked_prism_IL) +
  tm_raster() +
tm_shape(IL_county) +
  tm_polygons(alpha = 0)
```







## Descriptive Statistical: ggpairs() {#CorrelationMap}

```{r}
ggpairs()

data(diamonds, package="ggplot2")
diamonds.samp <- diamonds[sample(1:dim(diamonds)[1], 200), ]

# Custom Example  ( almost directly from help page)
pm <- ggpairs(
 diamonds.samp[, 1:5],
 mapping = ggplot2::aes(color = cut),
 upper = list(continuous = wrap("density", alpha = 0.5), combo = "box"),
 lower = list(continuous = wrap("points", alpha = 0.3,    size=0.1), 
              combo = wrap("dot", alpha = 0.4,            size=0.2) ),
 title = "Diamonds"
)
```

## theme() of ggplot2 {#theme}

```{r}
theme_figure <-
  theme(
    plot.title = element_text(hjust = 0.5),
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks = element_blank(),
    rect = element_blank(),
    text = element_text(family = "Times New Roman")
  )

theme_dist <-
  theme_bw() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    strip.text.x = element_text(size = 12, face = "bold"),
    strip.text.y = element_text(size = 12, face = "bold"),
    legend.title = element_text(size = 12, face = "bold"),
    legend.text = element_text(size = 12, face = "bold"),
    legend.position = "bottom",
    text = element_text(family = "Times New Roman")
  )
```

### How to increase the size of points in legend of ggplot2?

```{r}
+ guides(colour = guide_legend(override.aes = list(size=10)))
```

### tweak a map to make it much nicer look

```{r}
g_base + theme(axis.text = element_blank())     

theme_for_map <- theme(
  axis.ticks = element_blank(),
  axis.text= element_blank(), 
  axis.line = element_blank(),
  panel.border = element_blank(),
  panel.grid = element_line(color='transparent'),
  panel.background = element_blank(),
  plot.background = element_rect(fill = "transparent",color='transparent')
)

g_base + theme_for_map
```

### Change ggplot facet lavels {#GgplotFacetLavels}

```{r}
library(ggplot2)
theme_set(
  theme_light() + theme(legend.position = "top")
  )

# Load data and convert dose to a factor variable
data("ToothGrowth")
ToothGrowth$dose <- as.factor(ToothGrowth$dose)

# Box plot, facet accordding to the variable dose and supp
p <- ggplot(ToothGrowth, aes(x = dose, y = len)) + 
  geom_boxplot(aes(fill = supp), position = position_dodge(0.9)) +
  scale_fill_viridis_d() 
p + facet_grid(dose ~ supp)


# New facet label names for dose variable
dose.labs <- c("D0.5", "D1", "D2")
names(dose.labs) <- c("0.5", "1", "2")

# New facet label names for supp variable
supp.labs <- c("Orange Juice", "Vitamin C")
names(supp.labs) <- c("OJ", "VC")

# Create the plot
p + facet_grid(
  dose ~ supp, 
  labeller = labeller(dose = dose.labs, supp = supp.labs)
  )
```



### Change legend position (1)

```{r}
g_legend + 
    scale_fill_viridis_c() + 
    labs(fill = "Groundwater pumping (acre-feet)") +
    theme(legend.position = "bottom")
```

### Change legend position (2)
+ have the legend title on top of the color bar with guides()

```{r}
g_legend + 
    scale_fill_viridis_c() + 
    labs(fill = "Groundwater pumping (acre-feet)") +
    theme(legend.position = "bottom") +
    guides(fill = guide_colorbar (title.position="top"))
```

### Elongate the color bar and make the legend thinner

with `legend.key.width =` option and `legend.key.height=` for `theme()`

```{r}
g_legend + 
    scale_fill_viridis_c() + 
    labs(fill = "Groundwater pumping (acre-feet)") +
    theme(
        legend.position = "bottom",
        #--- NEW LINES HERE!! ---#
        legend.key.height = unit(0.5, "cm"),
        legend.key.width = unit(2, "cm")
    ) +
    guides(fill = guide_colorbar(title.position="top"))
```

<br>

**In the case of discrete values, use `guide_legend()` in the place of `guide_colorbar` of `guide()` function**

```{r}
g_legend_2 <- ggplot(data = gw_Stevens) +   
    geom_sf(aes(color = af_used_cat), size = 2) +
    scale_color_viridis(discrete = TRUE, option = "C") +
    labs(color = "Groundwater pumping (acre-feet)")  +
    theme_void() +
    theme(legend.position = "bottom") 

g_legend_2 +
    guides(
        color = guide_legend(title.position="top") 
    )
```


## Modify the legend {#ChangeLegendName}

+ you can change the legend name with `scale_fill_viridis_c()`
+ Or, use `labs()`

```{r}
g_legend +
    scale_fill_viridis_c(name = "Groundwater pumping (acre-feet)")  

g_legend +
    scale_fill_viridis_c() + 
    labs(fill = "Groundwater pumping (acre-feet)") 

# Or 
guides(fill=guide_legend(title="New Legend Title"))
```

+ Remove a legend

```{r}
theme(legend.position="none")
```

+ Remove a legend title

```{r}
theme(legend.title=element_blank())
```

+ Modify the label of the legend: `scale_fill_hue` or `scale_color_hue`
  * Especially in the case when you want to change the group names of the legend.
  * Other options are `scale_color_discrete` or `scale_fill_discrete`

```{r}
ggplot() +
  geom_sf(data=nrd_boud)+
  geom_sf(data=st_as_sf(ir_data, coords = c("longdd","latdd"), crs = 4269), aes(color=factor(in_5mi)), size=0.05) +
  theme_bw() + 
    theme(
    plot.title = element_text(hjust = 0.5),
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks = element_blank()
  ) +
  scale_color_hue(labels = c("Wells inside of the buffer", "Wells outside of the buffer"))
  ggtitle("(1) Wells within the 5 mile buffer")

```



## Plot Raster with facet {#RasterFascet}

### `ggplot2`

```{r}
test <-  
  as.data.frame(pr_all, xy = TRUE) %>%
  melt(., id.vars = c("x", "y"))

 ggplot() +
    geom_sf(
      data = 
        nrd_bound %>%
        filter(NRD_Name == "Lower Republican") %>%
        st_transform(crop_layer, crs = terra::crs(temp_crop_in_inch_sum))
      ) +
    geom_raster(data = test, aes(x = x, y = y, fill = value)) +
    facet_wrap(~ variable) +
    scale_fill_viridis_c() +
    theme(
      legend.position = "bottom"
    ) +
    labs(fill = "In-season precipitation (inches)") 
```


### `rasterVis`

```{r}
library(rasterVis)
vis_pr <- 
  gplot(pr_all)  +
    geom_tile(aes(fill = value)) +
    facet_wrap(~ variable) +
    scale_fill_gradientn(colours = rev(terrain.colors(225))) +
    theme(
      legend.position = "bottom"
    ) +
    labs(fill = "In-season precipitation (inches)") 
```

## Color command {#color}
```{r}
scale_fill_brewer(palette = "Paired")
```

```{r}
ggplot(bd_nrd_for_map) +
  # === Basic info === #
  geom_sf(fill=NA) +
  # === years === #
  geom_sf(aes(fill=Year), alpha=0.8)+
  scale_fill_manual(values=my_cols)+
  # === NRD name === #
  geom_text(aes(x = X, y = Y, label = NRD_Name), size = 3)+
  # === format === #
  theme_void() +
  labs(
    title = "The implementation of groundwater allocation policies by year",
    fill = "First Implementation Year"
  ) +
  theme(
    plot.title = element_text(hjust = 0.5)
  )

# === Set fill colors === #
alloc_colors <- c(
  # create colors for year variables (6 unique years)
  brewer_pal("div", palette = "Greens", direction = -1)(6),
  # colors for non-year variables
  c("darkgray", "orange")
  )


ggplot(bd_nrd_for_map) +
  # === Basic info === #
  geom_sf(fill=NA) +
  # === years === #
  geom_sf(aes(fill=alloc_year), alpha=0.8)+
  scale_fill_manual(values=alloc_colors)+
  # === NRD name === #
  geom_text(aes(x = X, y = Y, label = NRD_Name), size = 3)+
  # === format === #
  theme_void() +
  labs(
    title = "The implementation of groundwater allocation policies by year",
    fill = "First Implementation Year"
  ) +
  theme(plot.title = element_text(hjust = 0.5))


  
tm_shape(bd_nrd_for_map) +
  tm_polygons(col="alloc_year", title = "First Implementation Year") +
  tm_text("NRD_Name", size=0.8) +
  tm_layout(
      main.title = "The implementation of groundwater allocation policies by year", 
      main.title.position = "center",
    frame = FALSE
    )
```


## ggplot with year {#year}

```{r}
all_stn_count_m <- 
  readRDS(here(base_dir, "z_Writing/all_stn_count_m.rds"))

fig_all_stn_count <- 
  ggplot(all_stn_count_m)+
  geom_line(aes(x=year_month, y=Count, group=var, color=var))+
  scale_x_date(
    date_breaks = "1 year", 
    # date_labels = "%b/%Y"
    date_labels = "%Y"    
  ) +
  theme_bw() +
  # theme(
  #   axis.text.x=element_text(angle=60, hjust=1))+
  labs(x = "Year-Month", y = "Count") +
  theme(
    legend.position = "bottom"
  ) +
  scale_color_discrete(
    name = "Weather stations",
    labels=c("Precipitation", "Maximum temperatures", "Minimum temperatures")
    )
```



---

# Useful Functions

## lab {#lab}

```{r}
labs(fill = "Groundwater pumping (acre-feet)")
labs(fill = “Yield level \n (kg/ha)”
```

## theme {#theme}

```{r}
theme(panel.grid = element_line(color="transparent"),  )

theme(legend.position = "bottom")

theme(
  legend.position = "bottom",
  #--- change the size of the legend ---#
  legend.key.height = unit(0.5, "cm"),
  legend.key.width = unit(2, "cm")
  )
```


## facet_grid(type~variable,scale='free') {#AdjustScaleAxis}

```{r}
g_impact_viz <- ggplot() +
    geom_line(data=impact_data, aes(y=tau_hat,x=value)) +
    geom_ribbon(data=impact_data,
      aes(
      #- 95% CI: lower bound -#
        ymin=tau_hat-1.96*tau_hat_se,
        #- 95% CI: upper bound -#
        ymax=tau_hat+1.96*tau_hat_se,
        x=value
      ),
      alpha=0.4
    ) +
    geom_histogram(data=data_dist_viz,aes(x=value),color='black',fill='white') +
    facet_grid(type~variable,scale='free') +
    ylab('') +
    xlab('')
```



## theme(panel.grid = element_line(color='transparent')) {#Transparent}

```{r}
ggplot() +  
    geom_sf(data = gw_by_county, aes(fill = af_used)) +
    facet_grid(year ~ .) + 
    theme_bw() +
    theme(
        panel.grid = element_line(color='transparent'))
    )
```


## scale_fill_viridis_c() {#scale_fill_viridis_c}

```{r}
g_legend +
    scale_fill_viridis_c(name = "Groundwater pumping (acre-feet)")  
```













