---
title: "Functions for GIS operation"
author: "Shunkei Kakimoto"
output:
  html_document:
    number_sections: yes
    theme: flatly
    toc_float: yes
    toc: yes
    toc_depth: 3
geometry: margin=1in
---

```{r setup, include=FALSE}
library(knitr)
library(here)

here::i_am("Dropbox/R-project/R-functions/GitControlled/GIS/R-GIS-raster-concept.rmd")

# opts_knit$set(root.dir = "")


knitr::opts_chunk$set(
  echo = FALSE,
  eval=FALSE,
  cache = FALSE,
  comment = NA,
  message = FALSE,
  warning = FALSE,
  tidy = FALSE,
  cache.lazy = FALSE,
  #--- figure ---#
  dpi = 400,
  fig.width = 7.5,
  fig.height = 5,
  out.width = "750px",
  out.height = "500px"
)

# /*===== Basic Packages  =====*/
# /*---- Data Wrangling ----*/
library(data.table)
library(tidyverse)
library(DescTools)
library(maps)

# /*---- Visualization ----*/
library(RColorBrewer)
library(patchwork)
library(ggplot2)
library(ggthemes)
library(ggpubr)
library(viridis)
library(grid)
library(gridExtra)
library(GGally)

# /*---- Model Summary ----*/
library(stats)
library(modelsummary)
library(flextable)
library(officer)
library(officedown)
library(gt)

```


# Raster

+ `raster` package
+ `terra` package
+ `stars` package
	* work with raster data with the temporal dimension (e.g., PRISM, Daymet)
	* It also allows you to apply dplyr verbs for data wrangling


## `raster` package

has multiple object classes depending on a single or multiple raster objects. 

+ `RasterLayer` class
	* consists of only one layer, meaning that only a single variable is associated with the cells

+ `RasterStack` class
	* `raster::stack()`
	* tack multiple raster layers

+ `RasterBrick` class
	* `raster::brick()`

```{r}
#--- stack the two ---#
IA_cdl_brick <- brick(IA_cdl_stack)

#--- or this works as well ---#
IA_cdl_brick <- brick(IA_cdl_2015, IA_cdl_2016)	
```	


## `terra` package

### Basic

has only one object class for raster data regardless of one-layer and multi-layer rasters.

+ `SpatRaster` class
	* `terra::rast()`
    -  this can convert a Raster class to a SpatRaster class object

If you want to stack multiple multiple single-layer `SpatRaster` objects, you can use `c()` like below.

```{r}
#--- create a single-layer SpatRaster ---#
IA_cdl_2015_sr <- rast(IA_cdl_2015)
IA_cdl_2016_sr <- rast(IA_cdl_2016)

#--- concatenate ---#
IA_cdl_ml_sr <- c(IA_cdl_2015_sr, IA_cdl_2016_sr)
```

Or you can import two single-layer raster datasets to create a two-layer `SpatRaster` object at the same time. 
+ NOTE: this only works because the two datasets have the identical spatial extent and resolution.

```{r}
#--- the list of path to the files ---#
files_list <- c("Data/IA_cdl_2015.tif", "Data/IA_cdl_2016.tif")

#--- read the two at the same time ---#
multi_layer_sr <- rast(files_list)
```

Use `terra::writeRaster()` to save `SpatRaster` object.

```{r}
terra::writeRaster(
	IA_cdl_2015_sr, 
	"Data/IA_cdl_stack.tif", 
	filetype = "GTiff", 
	overwrite = TRUE)	
```	

### Extract information from raster data object

+ `terra` package has its own class for vector data, called `SpatVector`. Whenever you use `SpatRaster` object with `sf` object, you need to convert the `sf` object to `SpatVector` object. 
	* `SpatVector` object
		- `terra::vect()`


+ `terra::crs()` to get CRS

+ `IA_cdl_stack_sr[[2]]` to subset a specific laysers in a multi-layer raster object 

+ ` terra::values()` to get cell values


+ Turning a raster object into a `data.frame` to create maps with `ggplot()`
	* ` as.data.frame(IA_cdl_stack_sr, xy = TRUE)`

```{r}
#--- converting to a data.frame ---#
IA_cdl_df <- as.data.frame(IA_cdl_stack_sr, xy = TRUE) # this works with Raster* objects as well
#--- take a look ---#
head(IA_cdl_df)
```

<br>

+ `terra::crop()`

```{r}
crop(raster object, geographic extent)
```














