---
title: "R Function Cheat Sheet"
author: "Shunkei Kakimoto"
output:
  html_document:
    number_sections: yes
    theme: flatly
    toc_float: yes
    toc: yes
    toc_depth: 2
geometry: margin=1in
---

```{r setup, include=FALSE}
library(knitr)
library(here)

here::i_am("Dropbox/R-project/R-functions/GitControlled/R-func-cheat-sheet.rmd")

# opts_knit$set(root.dir = "")
# opts_knit$set(root.dir = here())

knitr::opts_chunk$set(
  echo = TRUE,
  cache = FALSE,
  eval = FALSE, 
  comment = NA,
  message = FALSE,
  warning = FALSE,
  tidy = FALSE,
  cache.lazy = FALSE,
  #--- figure ---#
  dpi = 400,
  fig.width = 7.5,
  fig.height = 5,
  out.width = "750px",
  out.height = "500px"
)

# /*===== Basic Packages  =====*/
# /*---- Data Wrangling ----*/
library(data.table)
library(tidyverse)
library(DescTools)
library(maps)

# /*---- Visualization ----*/
library(RColorBrewer)
library(patchwork)
library(ggplot2)
library(ggthemes)
library(ggpubr)
library(viridis)
library(grid)
library(gridExtra)
library(GGally)

# /*---- Model Summary ----*/
library(stats)
library(modelsummary)
library(flextable)
library(officer)
library(officedown)
library(gt)

```

# Contents

## Topic
+ [Source all the functions in the Functions folder](#SourceFunction)
+ [Data Table Operation](#data_table)
+ [Manage strings](#ManageStrings)
+ [Basic Statics Summary](#SummaryTable)


## A list of useful functions

### High
+ [**melt()**](#melt)
	* wide- to long-format data

+ [**dcast()**](#dcast)
	* long- to wide-format data 

+ [**download()**](#Downlowd): Downlaod files: downloader package

+ <span style="color:green">**na.omit(., cols = "usage")**</span>
	* remove NA from specific columns

+ [**as.formula()**](#formula)

+ [**eval()**](#eval)
	* evaluate an R expression in a specified environment

+ <span style="color:green">**as.numeric(as.vector(temp.df_for4[i,]))**</span>
	* * convert a list data to vector
+ <span style="color:green">**do.call(c, lapply(seq(to=2241, by=160), create.number_for4))**</span>x
	* stratified list data to a single list data

+ <span style="color:green">**unlist()**</span>
		-  this is useful when using lapply function

+ [***stringr***](#stringr)
	* <span style="color:green">str_detect()</span>
	* <span style="color:gren">str_pad(string, width, side = c("left", "right", "both"), pad = " ")</span>

+ <span style="color:green">**fread()**</span>: when you read a large file (ex. .txt)

<br>

### Low
+ <span style="color:blue">setcolorder()</span>
	* change the column order

+ <span style="color:blue">conv_unit(60, "ft", "m")</span>
	* Converts common units of measurement for a variety of dimensions

+ <span style="color:blue"> </span>

+ <span style="color:blue">tempfile()</span>

+ [**list.files()**](#listfiles)
	* produce charactervector of the names of files or directories in the named directory

+ <span style="color:blue">seq_length()</span>
	* e.g., seq_length(4) produces 1,2,3,4 

+ [**while loop**](#whileloop)

+ <span style="color:blue">require()</span>

+ <span style="color:blue">arrange()</span>
	*  ascending order
	* <span style="color:blue">arrange(, desc())</span>
		-  descending order

+ [**shift()**](#shift)
	* calculate lead/lag for cectors and lists

+ <span style="color:blue">nchar()</span>
	* character counting 

+ <span style="color:blue">summarise()</span>
	* is a convenient function to create a quick data frame

---

# Topics

## Data Table Operation {#data_table}
+ `DT[[ ]]`
	* substruct single column from data.table as a vector using the name of the column

+ `rbindlist()`: combine a list of of the datasets into a single data.table
+ `rbindlist(tmax_by_county, idcol = "id")`   

+ `CJ( )`: from `data.table` package, (C)ross (J)oin. A data.table is formed from the cross product of the vectors. For example, 10 ids, and 100 dates, CJ returns a 1000 row table containing all the dates for all the ids. It gains sorted, which by default is TRUE for backwards compatibility. FALSE retains input order.

+ `fifelse()`: from `data.table` package. A faster and more robust replacement of `ifelse` fucntion.
	* the same syntax as `ifelse()` function
	* `fifelse(test, yes, no, na=NA)`

+ `data.table(, roll=TRUE)`: forward par of the rolling join

```{r}
data.table(, roll=TRUE): (f)
dt2[dt1, roll=TRUE, on=“joint_y”]
```

for each joint_y of dt1, finds joint_y of dt2 that is closest to joint_y of dt1 with the condition joint_y of dt1 >= joint_y of dt2

<br>

+ If you want to select columns with a sequence of coolumn names

```{r}
report_res_allML[,select(.SD, Method, Model, Min.:Max., cumulative_pi_loss)]
```


```{r}
CJ();
N <- 1000
T <- 10

reg_data <- 
  CJ(id = 1:N, t = 1:T) %>% 
  #=== individual FE ===#
  .[, ind_fe := rnorm(1), by = id] %>% 
  #=== year FE ===#
  .[, time_fe := rnorm(1), by = t] %>% 
  #=== covariates (independent) ===#
  .[, 
    `:=`(
      x_1 = rnorm(1), 
      x_2 = rnorm(1),
      x_3 = rnorm(1)
    ), 
    by = .(id, t)
  ]
```

<br>

## Source all the functions in the Functions folder {#SourceFunction}

```{r}
fs::dir_ls(here("Codes", "Functions", "R"), full.names = TRUE) %>%
  lapply(., function(x) source(x))
```

## Manage strings {#ManageStrings}

+ `grep()`: search for matches to argument `pattern` within each element of a character vectore

```{r}
new_names[#]
```

<br>

+ `gsub()`:
	* perform replacement of detail and all matches respectively remove all text before some symbol: 
	* `gsub(pattern, replacement, x, ignore.case = FALSE, perl = FALSE, fixed = FALSE, useBytes = FALSE)`


```{r}
# this remove all texts before “=“
gsub(".*=","","air_temperature_day=40446") 

text <- "40015"
# text <- "6047"
x <- sub("(.*)(.{3})", "\\1_\\2", text)
gsub("_.*", "", x) 
gsub(".*_", "", x) 


gsub(“20”, “”, as.character(2000:2019))


# replace “ “(space) with “_”(undervar)
new_names <- names(subset_prop_full) %>%
	 gsub(pattern = " ", replace="_") %>%
	 tolower()
```

+ `substr()`

```{r}
substr("abcdef", 2, 4)
```

<br>

+ `str_extract()`: extract matching patterns from a string
+ `str_extract(string, pattern)`

```{r}
shopping_list <- c("apples x4", "bag of flour", "bag of sugar", "milk x2")

str_extract(variable,"[0-9]{1,5}")] 
```

<br>

+ `str_pad()`: pad a string 

<br>

+ `str_sub`: extract and replace rebstrings from a character vector

```{r}
x <- "some text in a string"
str_sub(x,-6,-1)
```

<br>

+ `str_split`: solit up a string into pieces


## Basic Statics Summary {#SummaryTable}

+ modelsummary
+ gt

```{r}
datasummary_skim(full_dt[,.(total_value_calculated, assessed_total_value, market_total_value, appraised_total_value, sale_amount)],
	 title = "Quick overview of key variables in Full-data (all states)")

eq1 <- as.formula(paste(paste0(weather_var, collapse = "+"), "nrdname * (Mean + SD)",  sep="~"))

reg_data_new[,c("nrdname", weather_var, soil_var), with = FALSE] %>%
	datasummary(
		eq1,
		 # pr_in+tmmn_in ~  nrdname * (Mean + SD),
    data = .
    )

#- for row striping -#
custom_theme <- function(x, ...) {
    x %>% gt::opt_row_striping(row_striping = TRUE)
}
options("modelsummary_theme_gt" = custom_theme)
```

---

# A list of useful functions

## melt() {#melt}
+ wide- to long-format data

```{r}
DT.m1 = melt(DT, id.vars = c("family_id", "age_mother"),
                measure.vars = c("dob_child1", "dob_child2", "dob_child3"))   

melt(id.vars = "ID", variable.name = "day")
```

+ **Note**: 
	* measure.vars specify the set of columns we would like to collapse together.
	* id.vars specifies the variables that we don't change
	* By default, variable column is of type factor. Set variable.factor argument to FALSE if you’d like to return a character vector instead.


## dcast() {#dcast}
+ long- to wide-format data 

```{r}
dcast(DT.m1, family_id + age_mother ~ child, value.var = "dob")
```

+ **Note**:
	* LHS represents the id vars 
	* RHS the measure vars
	* value.var denotes the column to be filled in with while casting to wide format
	-> id vars ~ measure vars



## formula()

```{r}
as.formula()

eq1 <- as.formula(paste(paste0(weather_var, collapse = "+"), "nrdname * (Mean + SD)",  sep="~"))
```

## eval() {#eval}

```{r}
eval(parse(text=paste('data[,treat:=',treat_var,']',sep="")))
```

## lapply vs sapply
+ `lapply()` returns outputs as a list
+ `sapply()`returns outputs as a vector

<br>

Suppose you create a function that returns multiple object as a list. Then, after all iterations, you'll get a final results as a list that contains multiple object within it. If you want to retrieve one of them, `sapply` function is useful.

```{r}
sapply(all_results,'[',1)
sapply(all_results,'[',2) %>% rbindlist()

cf <- sapply(all_results,'[',1)
tau_all <- sapply(all_results,'[',2) %>% rbindlist()
vi_all <- sapply(all_results,'[',3) %>% rbindlist() %>% 
	.[,variable:=factor(variable,levels=var_ls)] %>% 
    dcast(variable~trt,value.var ='vi')
```


## download() {#Downlowd}

+ `download()` from `downloader` package

```{r}
download( url = , destfile = , mode = ‘web’ )
```

<br>


## stringr {#stringr}

+ `str_detect()`
	* Package "stringr"
	* Vectorised over string and pattern.
	* `str_detect(string, pattern, negate=FALSE)`

```{r}
tif_file_names <- list.files(path="./Data/HPA_sat_thickness_inputs" , pattern="*.tif", full.names=TRUE) %>% 
  .[!str_detect(.,'tif.')]

fruit <- c("apple", "banana", "pear", "pinapple")
str_detect(fruit, "a")       #ans. [1] TRUE TRUE TRUE TRUE
str_detect(fruit, "^a")      #ans. [1]  TRUE FALSE FALSE FALSE
str_detect(fruit, "a$")      #ans. [1] FALSE  TRUE FALSE FALSE
str_detect(fruit, "b")       #ans. [1] FALSE  TRUE FALSE FALSE
str_detect(fruit, "[aeiou]") #ans. [1] TRUE TRUE TRUE TRUE
```

+ `^` : now matches the start of each line
+ `$` : now matches the end of each line
+ `\A` : matches the start of the input
+ `\z` : matches the end of the input
+ `\Z` : matches the end of the input, but before the final line terminator, if exists



## list.files() {#listfiles}

produce charactervector of the names of files or directories in the named directory

```{r}
tif_file_names <- list.files(path="./Data/HPA_sat_thickness_inputs" , pattern="*.tif", full.names=TRUE) %>% 
  .[!str_detect(.,'tif.')]
```


## While loop {#whileloop}

A loop is a statement function that keeps running until a condition is satisfied.

+ usage: 

```{r}
while (condition){
                Exp
        }
```

For exampel: 

```{r}
total <- 0
    number <- as.integer(readline(prompt="Please Enter any integer Value below 10:  "))

    whiel(number <= 10) {
    total = total + number
    number = number + 1
    }
```


## shift() {#shift}

```{r}
.[, tau_stepwise := yield_hat - shift(yield_hat), by = unique_cell_id]
```








