---
title: "R Function Cheat Sheet"
author: "Shunkei Kakimoto"
output:
  html_document:
    number_sections: yes
    theme: flatly
    toc_float: yes
    toc: yes
    toc_depth: 2
geometry: margin=1in
---

```{r setup, include=FALSE}
library(knitr)
library(here)

here::i_am("Dropbox/R-project/R-functions/GitControlled/General/R-func-cheat-sheet.rmd")

# opts_knit$set(root.dir = "")
# opts_knit$set(root.dir = here())

knitr::opts_chunk$set(
  echo = TRUE,
  cache = FALSE,
  eval = FALSE, 
  comment = NA,
  message = FALSE,
  warning = FALSE,
  tidy = FALSE,
  cache.lazy = FALSE,
  #--- figure ---#
  dpi = 400,
  fig.width = 7.5,
  fig.height = 5,
  out.width = "750px",
  out.height = "500px"
)

# /*===== Basic Packages  =====*/
# /*---- Data Wrangling ----*/
library(data.table)
library(tidyverse)
library(DescTools)
library(maps)

# /*---- Visualization ----*/
library(RColorBrewer)
library(patchwork)
library(ggplot2)
library(ggthemes)
library(ggpubr)
library(viridis)
library(grid)
library(gridExtra)
library(GGally)

# /*---- Model Summary ----*/
library(stats)
library(modelsummary)
library(flextable)
library(officer)
library(officedown)
library(gt)

```

+ `tidycensus`: 
  * Details [here](https://walker-data.com/census-r/an-introduction-to-tidycensus.html)
  * Functions [here](https://walker-data.com/tidycensus/index.html)

+ NOTE: `tidycensus` does not grant access to all of the datasets available from the Census API

# Main

## **Decennial US Census**
+ Every 10 years: 2000, 2010, 2020

```{r}
# Ex)
get_decennial(
  geography = "state", 
  variables = "P001001",
  year = 2010,
  sumfile = "sf1" #(by default)
)
```

+ The Census summary file
  * `sumfile = "sf1"`: the decennial Census Summary File 1, and includes core demographic characteristics for Census geographies
  * `sumfile = "sf2"`: the decennial Census Summary File 2, contains information on a range of population and housing unit characteristics
  * "sf3" and "sf4": Detailed demographic information in the 2000 decennial Census such as income and occupation can be found
  * `sumfile = "pl"`: for 2020 Decennial Census data. 

+ 2000 Census data: "sf1", "sf2", "sf3" and "sf4"
+ 2010 Census data: "sf1" and "sf2"
+ 2020 Census data: use "pl" (the only dataset supported for now)

+ The smallest unit at which available from the decennial US Census is the "block"
  * * the boundary is available at `tigris::blocks(state, county = NULL, year = NULL, ...)`


## **American Community Survey (ACS)**:
+ the premier source of detailed demographic information about the US population.
+ Two official datasets:
  * 1-year ACS: covers areas of population 65,000 and greater (from 2005)
  * 5-year ACS: a moving average of data over a 5-year period that covers geographies down to the Census block group (from 2005-2009)

+ *NOTE*: ACS data are distinct from decennial Census data in that **data represent estimates rather than precise counts**, and in turn are characterized by margins of error around those estimates


```{r}
# Ex)
# for 5-year ACS fof 2016-2020
get_acs(
  geography = "state", 
  variables = "B05006_150",
  survey = "acs5", # by default
  year = 2020
  
)

# for 1-year ACS for 2019
 get_acs(
  geography = "state", 
  variables = "B05006_150", 
  survey = "acs1", # change to "acs1" to specify year
  year = 2019
)

```


<br>

+ The smallest unit in the ACS is the "block" group (> Census block)
  * the boundary is available at `tigris::block_groups()`





## How to search Census variable ID?

+ `load_variables(year, dataset)`: obtains a dataset of variables from the Census Bureau website and formats it for fast searching, ideally in RStudio.



# Other Census Bureau datasets in tidycensus

## get_estimates()
+ Population Estimates Program (PEP) provides **yearly** estimates of the US population and its components between decennial Censuses.
  * PEP is derived from the most recent decennial Census based on birth, death, and migration rates. 
  * One advantage of using the PEP to retrieve data is that allows you to access the indicators used to produce the **intercensal population estimates**.
    - `product`
      + "population"
      + "comments"
      + "housing"
      + "characteristics"

```{r}
queens_components <- get_estimates(
  geography = "county",
  product = "components",
  state = "NY",
  county = "Queens",
  year = 2019
)
```






```{r}
# /*===== Functions =====*/
dissove_bd_by_county <- function(data, state){
  # data=hp_urban_blocks ; state="Nebraska"

  temp_data <- filter(data, State==state)

  return_res <- 
    tibble(
      section = unique(temp_data$County)
    ) %>%
    rowwise() %>%
    mutate(
      section_dt = list(
        filter(temp_data, County == section)
      )
    ) %>%
    ungroup() %>%
    mutate(
      agg_section_dt = future_lapply(
        seq_len(nrow(.)),
        function(x){
          .$section_dt[[x]] %>%
            group_by(County, State) %>%
            summarise()
        }
      )
    ) %>%
    unnest(., cols= "agg_section_dt") %>%
    mutate(
      section = NULL,
      section_dt = NULL,
    ) %>%
    st_as_sf()

    return(return_res)
}


# /*===== Apply to each state  =====*/
tic()
res_urban_dissolve <- 
  lapply(
    unique(hp_urban_blocks$State),
    function(x) 
      dissove_bd_by_county(
        data = hp_urban_blocks,
        state = x)
      ) %>%
  bind_rows()
toc()
```















