---
title: "R GIS Function Cheat Sheet"
author: "Shunkei Kakimoto"
output:
  html_document:
    number_sections: yes
    theme: flatly
    toc_float: yes
    toc: yes
    toc_depth: 3
geometry: margin=1in
---

```{r setup, include=FALSE}
library(knitr)
library(here)

here::i_am("Dropbox/R-project/R-functions/GitControlled/GIS/R-GIS-general.rmd")

# opts_knit$set(root.dir = "")
# opts_knit$set(root.dir = here())

knitr::opts_chunk$set(
  echo = TRUE,
  eval =FALSE,
  cache = FALSE,
  comment = NA,
  message = FALSE,
  warning = FALSE,
  tidy = FALSE,
  cache.lazy = FALSE,
  #--- figure ---#
  dpi = 400,
  fig.width = 7.5,
  fig.height = 5,
  out.width = "750px",
  out.height = "500px"
)

# /*===== Basic Packages  =====*/
# /*---- Data Wrangling ----*/
library(data.table)
library(tidyverse)
library(DescTools)
library(maps)

# /*---- Visualization ----*/
library(RColorBrewer)
library(patchwork)
library(ggplot2)
library(ggthemes)
library(ggpubr)
library(viridis)
library(grid)
library(gridExtra)
library(GGally)

# /*---- Model Summary ----*/
library(stats)
library(modelsummary)
library(flextable)
library(officer)
library(officedown)
library(gt)

```

# Content

## Topics
+ [Download state, county border](#DonloadStateBoundary)
+ [Make boundary sf object from multi-polygons](#MakeBoundary)
+ [Create X, Y columns](#CreateXYColumns)
+ [Make sf object from data.table](#Makesf)
+ [Spatial Operations on vector data](#SpatialOperationsVector)
  * See "Geocomputation with R":[here](https://geocompr.robinlovelace.net/spatial-operations.html)
  * 1. Spatial subsetting
  * 2. `ifelse` with spatial condition


## Useful Functions

+ [**map()**](#map)
  * Draw lines and polygons as specified by a map database

+ [**values()**](#value)
  * returns the raster cell values for a RasterLayer or RasterBrick object that are currently in memory

+ `tidyusda` package to get stete_fp - state name correspondence

+ <span style="color:gren">**nlayers()**</span>
  * returns the number of layers stored in a *Raster* object
+ <span style="color:green">**st_coordinate()**</span>
  * retrieve coordinates in matrix form
 
+ <span style="color:green">**raster::stack()**</span>

+ <span style="color:green">**raster::brick()**</span>

+ <span style="color:green">**terra::rast()**</span>
  * convert a Raster* to a SpatRaster

+ <span style="color:green">**terra::crop()**</span>
  * crop a RasterLayer 

+ <span style="color:green">**terra::ext()**</span>
  * find the geographical extent of a vector data. This case is happen when I want to use that extent to crop a RasterLayer.

+ <span style="color:green">**terra::extract(SpatRaster, SpatVector)**</span>
  * extract values from raster layers to points (from points)

+ <span style="color:green">**exact_extract(raster, sf)**</span>
  * extract values for polygons, (from the exactextractr package)

---

# Topics

## Download state, county border {#DonloadStateBoundary}

+ `tigris` package 

```{r}
state <- tigris::states(, cb = TRUE)

KS_county_sf <-
  #--- get Kansas county boundary ---
  tigris::counties(state = "Kansas", cb = TRUE) %>%
  #--- sp to sf ---#
  st_as_sf() %>%
  #--- transform using the CRS of the PRISM tmax data  ---#
  st_transform(terra::crs(prism_tmax_0701_sr))
```

<br>

<br>

## Make boundary sf object from multi-polygons {#MakeBoundary}

```{r}
field_base %>%
  st_bbox() %>%
  st_as_sfc() %>%
  st_as_sf()
## this might be better
st_union()
```

<br>

<br>


## Create X, Y columns {#CreateXYColumns}

```{r}
xy <- dplyr::select(field_padding, unique_cell_id) %>%
  cbind(., st_coordinates(st_centroid(.))) %>%
  st_drop_geometry() %>%
  data.table()
```

<br>

<br>

## Make sf object from data.table {#Makesf}

```{r}
st_as_sf(
  x =,
  coords = c("X", "Y"),
  crs = 
  )
```

<br>

<br>

## Spatial Operations on vector data {#SpatialOperationsVector}

### 1. Spatial subsetting
syntax `x[y, , op = st_intersects]`:
+  `x` is an sf object from which a subset of rows will be returned
  *  point 
+  `y` is the ‘subsetting object
  *  polygon

```{r}
canterbury_height = nz_height[canterbury, ]
```


### 2. `ifelse` with spatial condition
Use `st_intersects` for the condition:
  + syntax `st_intersects(x, y)`:

```{r}
tg_corelogic_sf <- 
  # --- Subset points intersecting the areas within 100 km boundary --- #
  corelogic_sf[hp_bd_buffer_100km,] %>%
  # --- Make a dummy "Ogallala" --- #
  mutate(Ogallala = 
    ifelse(st_intersects(., hp_bd, sparse = FALSE), 1, 0)
    )
```




---

# Useful Functions

## maps {#map}

+ `map()`: Draw lines and polygons as specified by a map database
  * `maps::map(database='', regions='',fill=fill=TRUE, plot=FALSE)`

```{r}
KS_county_sf <- st_as_sf(maps::map('state', fill=TRUE, plot=FALSE))
```






















